
\setstretch{1.4}
\section*{Preface}




This Master's Thesis details work completed during the spring of 2025 at École Polytechnique Fédérale de Lausanne (EPFL), as the final step towards the completion of the five year Cybernetics and Robotics Master'ss program with a specialization in Biomedical Cybernetics at the Norwegian University of Science and Technology (NTNU). The project was conducted in collaboration with the MICROBS Lab at EPFL in Lausanne, Switzerland, under the supervision of Øyvind Stavdahl from NTNU, with additional supervision at EPFL by Lorenzo Noseda and Prof. Selman Sakar. This work builds upon the PhD research of Lorenzo Noseda and forms part of the ongoing efforts at MICROBS to develop a new generation of flexible robotic devices for minimally invasive brain surgery.

First of all, I would like to thank Lorenzo Noseda for all the valuable discussions, the many days spent on device manufacturing, and, above all, for an enjoyable and close collaboration through both challenges and successes. I am also grateful to Selman Sakar for the opportunity to work on this project, and to everyone at the MICROBS Lab for creating such a welcoming and enjoyable environment over the past five months. A special thanks goes to my NTNU supervisor Øyvind Stavdahl for his enthusiasm, encouragement and thoughtful guidance, which have shaped both this project and my own growth as an engineer. Finally, my deepest gratitude goes to my family, friends, and partner for their unwavering and invaluable love and support throughout the years.




\setstretch{1.4}
